---
title: "Assignment1"
author: "Shone George Kutty Renjan"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

| cs     | Col2    | Col3ddssd |
|--------|---------|-----------|
| wfww,l | sfsvs;  | sdvdsv    |
| fsdcd  | dvdvsd; | sdvdssv   |
| wfwef  | fefvs   | dsvdsv    |

---

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Questions

## Question 1

```{r}
Course_grading <- data.frame(
  Course_Elements = c("Assignments", "Laboratories", "Projects", "In-Class Quizzes", "Quizzes (Online)", "Mid-Term Exams", "Final Exam"),
  Count = c("4", "16", "2", "TDH", "TDH", "2", "1"),
  Weight_Each = c("3.75%", "0.625%", "3.33% & 6.67%", "-", "-", "10%", "30%"),
  Total_Weight = c("15%", "10%", "10%", "10%", "5%", "20%", "30%")
)

print("COURSE DESCRIPTION")
print(Course_grading)


```

Total = 100% There is no curve in this course.

```{r}
grading_scale <- data.frame(
  Grade = c("A", "B", "C", "D", "F"),
  Percentage = c("90% - above", "80% - 89%", "60% - 79%", "50% - 59%", "Below - 50%")
)
grading_scale
```

## Question 2

(A) Make the coplot as the biologist required Hint: Use coplot(), Lab 1, the code provided, and plotting options pch and col to differentiate the MILE variable.

```{r}
ddt <- read.csv("DDT-1.csv")
m=with(ddt, as.numeric(factor(MILE))) # A

length(unique(m)) #B

coplot(LENGTH ~ WEIGHT | RIVER * SPECIES, data = ddt, col = unique(m))
```

(B) Interpret the lower left three conditional plots.

It shows a positive correlation between the length and weight among the CCATFISH species in all four rivers (FCM, LCM, SCM, TRM)

(C) What does line A do?

Line A creates a numeric vector 'm' from the 'MILE' column in the 'ddt' data frame. It first converts 'MILE' to a factor and then to numeric valuesâ 

(D) What does line B do?

Calculates the number of unique values in 'm' which is MILE

(E) Why are the top six plots empty?

The top six plots empty because there no sample for the species LMBASS and SMBUFFALO.

(F) What is the mean value of DDT found in the sample of CCATFISH caught in the FCM river?

```{r}
CCATFISH= ddt$DDT[ddt$SPECIES== "CCATFISH" & ddt$RIVER== "FCM"]
mean(CCATFISH)
```

## Question 3

### MS1.14

(a) Length of the maximum span (feet): quantitative
(b) Number of vehicle in lane: quantitative
(c) Toll bridge (yes or no): qualitative
(d) Average daily traffic: quantitative
(e) Condition of the deck (goof, fair, or poor): qualitative
(f) Bypass or detour length (miles): quantitative
(g) Route type(interstate, U.S., state, country, or city): qualitative

## Question 4

### MS

(A) What are the names of the four random sampling designs

The four random sampling designs are simple random sampling, stratified random sampling, Cluster Sampling, and Systematic Sampling.

(B) Give a brief description of each.

Simple Random Sampling: This method involves selecting a sample of size n from the population in such a way that every possible sample of size n has an equal chance of being selected.

Stratified Random Sampling: This method is used when the population can be divided into two or more distinct groups, or strata. A random sample is then taken from each group.

Systematic Sampling: This method involves selecting every k-th experimental unit from a list of all units in the population, starting from a randomly chosen point.

Cluster Sampling: In this method, the population is divided into natural groups, or clusters. Then, entire clusters are randomly selected, and data is collected from all units within these selected clusters.

## Question 5

### MS 1.15

Reads the MTBE File

```{r}
mtbe=read.csv("MTBE.csv", header=TRUE) # You will need to change the address
head(mtbe) # First six lines
dim(mtbe) # rows and columns
ind=sample(1:223,5,replace=FALSE) # random indices
mtbe[ind,]
```

(A) 

```{r}
mtbe=na.omit(mtbe)
depth=mtbe[mtbe$Aquifier=="Bedrock",]$Depth
sd(depth)

```

## Question 6

```{r}
# random sample 
earthquake = read.csv("EARTHQUAKE.csv")
earthquake[sample(1:nrow(earthquake), 30, replace=FALSE),]
```

(A) Make the following plot plot

```{r}
plot(ts(earthquake$MAG))
```

(B) Using earthquake data frame find the median

```{r}
median(earthquake$MAGNITUDE)
```

## Question 7

(A) What is the data collection method? The data was collection method was stratified sampling

(B) What is the population? Population fish in the Tennessee river.

(C) Give the names of all the qualitative variables

Qualitative variables: River and Species

## Question 8

(A) What type of graph is used to describe this data

Bargraph

(B) Identify the variable measured for each of the 106 robotic design

Robotic Limbs

(C) Use graph to identify the robot design currently most used

Legs

(D) Compute the relative frequencies for the most categories shown in the graph

```{r}
freq = c(15, 8, 63, 20)                      
fsum = sum(freq)
relfreq = freq / fsum
relativefrequencies <- data.frame(
  RL = c("None", "Both", "LegsO", "WheelsO"),
  relfreq
)
relativefrequencies

```

(e) pareto diagram

```{r}
pareto <- function(x, mn = "Pareto barplot", ...) {  # x is a vector
  x.tab = table(x)
  xx.tab = sort(x.tab, decreasing = TRUE, index.return = FALSE)
  cumsum(as.vector(xx.tab)) -> cs
  length(x.tab) -> lenx
  bp <- barplot(xx.tab, ylim = c(0,max(cs)),las = 2)
  lb <- seq(0,cs[lenx], l = 11)
  axis(side = 4, at = lb, labels = paste(seq(0, 100, length = 11), "%", sep = ""), las = 1, line = -1, col = "Blue", col.axis = "Red")
  for(i in 1:(lenx-1)){
    segments(bp[i], cs[i], bp[i+1], cs[i+1], col = i, lwd = 2)
  }
  title(main = mn, ...)
}
RL = c("None", "Both", "LegsO", "WheelsO")
l=rep(RL,freq)
pareto(l)

```

## Question 15

BOXPLOT

```{r}

```

| Col1 | Col2 | Col3 |
|------|------|------|
|      |      |      |
|      |      |      |
|      |      |      |

```{library(ggplot2)}
gg_plot <- ggplot(ddt, aes(x = RIVER, y = LENGTH, fill = SPECIES)) + geom_boxplot() + ggtitle("Shone George Kutty Renjan")
gg_plot
```
