---
title: "R Package"
author: "Shone George Kutty Renjan"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction
Congratulations for getting this far. To complete the evidence that you have made a working R package please answer the following by using the supplied rmd and upload the knitted html to CANVAS.

I will assume your package is called $MATH4753SHON2024$ and you have packaged and documented functions $mybin$, $cltp$, $mymaxlik$, and $bootpval$.$\\$
Your actual package name and functions will obviously be called something else.



# First Function
Please place your fourth function in the r chunk in the following way:

```{r}
MATH4753SHON2024::mybin()
```
# Second Function 
Please place your fourth function in the r chunk in the following way:

```{r}
library(MATH4753SHON2024)
graph<- cltp(n=7, iter = 10000)
```

# Third Function 
Please place your fourth function in the r chunk in the following way:

```{r}
logbin=function(x,param) log(dbinom(x,prob=param,size=20))

MATH4753SHON2024::mymaxlik(x=c(4,2,4,1,4,5,5,3),param=seq(0,1,length=1000),lfun=logbin,xlab=expression(pi),main="Binomial",cex.main=2)
```

# Fourth funtion
Please place your fourth function in the r chunk in the following way:

```{r}
set.seed(55); x1=rnorm(30,mean = 25,sd=5)
MATH4753SHON2024::bootpval(x1,mu=25)
```

This is the address of my GITHUB repository: https://github.com/Shonegr28/Rstudio
Replace this with your own address.

