---
title: "lab11"
author: "Shone George Kutty Renjan"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Tasks

## Task 1
```{r}
getwd()
```


## Task 2

Confidence Interval for mu:
$$\mu=y'+t_\frac{\alpha}{2}\frac{s}{\sqrt{n}}$$
```{r}
d=c(5.0581, 4.9707, 5.0893, 4.9334, 4.9777, 5.0285, 4.8555, 4.9565, 
4.9769, 4.9722, 4.999, 4.9925, 4.9686, 5.0662, 4.9239, 4.9781, 
5.0485, 5.0014, 4.9957, 5.0195, 5.0118, 4.9928, 5.0361, 5.0185, 
4.9879)
t=qt(0.975,24)
ci=c()
ci[1]=mean(d)-t*sd(d)/sqrt(25)
ci[2]=mean(d)+t*sd(d)/sqrt(25)
ci

```

```{r}
t=qt(0.95,24)
ci=c()
ci[1]=mean(d)-t*sd(d)/sqrt(25)
ci[2]=mean(d)+t*sd(d)/sqrt(25)
ci
```

```{r}

t=qt(0.9,24)
ci=c()
ci[1]=mean(d)-t*sd(d)/sqrt(25)
ci[2]=mean(d)+t*sd(d)/sqrt(25)
ci
```

```{r}

t=qt(0.75,24)
ci=c()
ci[1]=mean(d)-t*sd(d)/sqrt(25)
ci[2]=mean(d)+t*sd(d)/sqrt(25)
ci
```

```{r}
obj=t.test(d,conf.level = 0.8)
obj$conf.int
```

Confidence interval sigma^2:
$$\sigma^2=[\frac{(n-1)s^2}{X_\frac{\alpha}{2}^2},\frac{(n-1)s^2}{X^2_{1-}{\frac{\alpha}{2}}}]$$
```{r}
n = length(d)
xsq1=qchisq(0.975,(n-1))
xsq2=qchisq(0.025,(n-1))
ci=c()
ci[1]=(n-1)*var(d)/xsq1
ci[2]=(n-1)*var(d)/xsq2
ci
```
```{r}
n = length(d)
xsq1=qchisq(0.95,(n-1))
xsq2=qchisq(0.5,(n-1))
ci=c()
ci[1]=(n-1)*var(d)/xsq1
ci[2]=(n-1)*var(d)/xsq2
ci
```

```{r}
n = length(d)
xsq1=qchisq(0.9,(n-1))
xsq2=qchisq(0.1,(n-1))
ci=c()
ci[1]=(n-1)*var(d)/xsq1
ci[2]=(n-1)*var(d)/xsq2
ci
```

```{r}
n = length(d)
xsq1=qchisq(0.75,(n-1))
xsq2=qchisq(0.25,(n-1))
ci=c()
ci[1]=(n-1)*var(d)/xsq1
ci[2]=(n-1)*var(d)/xsq2
ci
```


## Task 3
```{r}
blue=c(21.65, 17.48, 20.1, 21.57, 14.82, 19.17, 21.08, 18.23, 22.93, 
15.66, 20.89, 21.66, 18.5, 20.59, 18.63, 18.91, 19.53, 17.7, 
16.5, 19.03)


snapper=c(31.65, 27.48, 30.1, 31.57, 24.82, 29.17, 31.08, 28.23, 32.93, 
25.66, 30.89, 31.66, 28.5, 30.59, 28.63)

n1=length(snapper)
n2=length(blue)
spsq=((n1-1)*var(snapper)+(n2-1)*var(blue))/(n1+n2-2)
t=qt(0.975,n1+n2-2)
ci=c()
ci[1]=mean(blue)-mean(snapper)-t*sqrt(spsq*(1/n1+1/n2)) 
ci[2]=mean(blue)-mean(snapper)+t*sqrt(spsq*(1/n1+1/n2))
ci
```

```{r}
n1=length(snapper)
n2=length(blue)
spsq=((n1-1)*var(snapper)+(n2-1)*var(blue))/(n1+n2-2)
t=qt(0.975,n1+n2-2)
ci=c()
ci[1]=mean(snapper)-mean(blue)-t*sqrt(spsq*(1/n1+1/n2)) 
ci[2]=mean(snapper)-mean(blue)+t*sqrt(spsq*(1/n1+1/n2))
ci
```
With a 95% confidence level, the Snappers' mean length was between 8.7576 and 11.8407 inches longer than the Blue Cod fish's mean length, according to the stated earlier interval.

```{r}
# 95% ci
t.test(snapper,blue,var.equal=TRUE,conf.level=0.95)$conf.int
```

```{r}
# 85% ci
t.test(snapper,blue,var.equal=TRUE,conf.level=0.85)$conf.int
```

```{r}
# 75% ci
t.test(snapper,blue,var.equal=TRUE,conf.level=0.75)$conf.int
```

```{r}
# 25% ci
t.test(snapper,blue,var.equal=TRUE,conf.level=0.25)$conf.int
```
As confidence decreases the interval gets smaller

## Task 4
```{r}
Exam1=c(40.98, 59.36, 46.69, 41.8, 61.63, 65.31, 62.96, 60.21, 56.89, 78.41, 53.44, 75.2, 60.54, 52.43, 41.41, 70.79, 73.55, 55.65, 61.43, 63.84, 58.07, 53.79, 54.45, 67.18, 44.46)
Exam2=c(50.22, 66.19, 58.75, 51.88, 66.61, 70.86, 74.25, 70.23, 69.55, 87.18, 63.62, 81.7, 70.5, 66.02, 51.35, 80.92, 85.65, 65.44, 74.37, 75.28, 67.86, 59.92, 64.42, 73.57, 57.15)
exdif=Exam1-Exam2
t.test(exdif,conf.level = 0.95)$conf.int
```
The mean score on Exam 1 is between 8.71 and 10.73 points lower than the mean score on Exam 2 with a 95% confidence level.
```{r}
t.test(exdif,conf.level=0.9)$conf.int

```
```{r}
t.test(exdif,conf.level=0.8)$conf.int

```

```{r}
t.test(exdif,conf.level=0.7)$conf.int

```

```{r}
t.test(exdif,conf.level=0.6)$conf.int

```

```{r}
t.test(exdif,conf.level=0.1)$conf.int

```


## Task 5
```{r}
df<- read.csv("NZBIRDS.csv")
with(df, table(Extinct,Flight))
```
```{r}
p_extinct_flightless = 7/(7+71)
p_extinct_flight= 21/(21+17)
z= qnorm(1-.05/2,0,1)
p_extinct_flightless*78 >=4
```

```{r}
p_extinct_flight*38>=4
```

```{r}
(1-p_extinct_flightless)*78>=4
```

```{r}
sq= sqrt((p_extinct_flightless*(1-p_extinct_flightless)/78)+(p_extinct_flight*(1-p_extinct_flight)/38))
pd = p_extinct_flight-p_extinct_flightless
c(pd-sq*z,pd+sq*z)
```

## Task 6
```{r}
set.seed(35);sam1=rnorm(25,mean=10,sd=5); set.seed(45);sam2=rnorm(34,mean=40,sd=8)
a=1-.05/2
ci=c()
ci[1]=var(sam1)/(var(sam2)*((qf(a, df1=24,df2=33))))
ci[2]=var(sam1)/var(sam2)*((qf(a, df1=33,df2=24)))
ci
```

```{r}
var.test(sam1,sam2,conf.level=0.8)$conf.int

```

```{r}
var.test(sam1,sam2,conf.level=0.7)$conf.int

```

```{r}
var.test(sam1,sam2,conf.level=0.6)$conf.int

```

```{r}
var.test(sam1,sam2,conf.level=0.5)$conf.int

```

## Task 7
```{r}
set.seed(23);x = rnorm(30,mean=10,sd=12)
MATH4753SHON2024::myci(x)
```

