---
title: "Assignment4"
author: "Jonathan Douglas"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
  pdf_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Set the WD

```{r}
setwd("/Users/jonathandouglas/MATH4753/Assignments/Assignment4/")
getwd()
```

## 1: 7.118

Read in the nzbirds data

```{r}
allData = Intro2R::myreadxl(dird = "../../DataForClass/Excel/")

bird_data = allData$NZBIRDS
```

#### Part A: Create random sample of birds

```{r}
#Use runif() to find the 35 numbers
species = bird_data$Species
set.seed(1)
sam = runif(35, 1, length(species))#Create list of random indicies
```

#### Part B: Find stats about the body mass of the birds

```{r}
sam_body_mass_full = bird_data$`Body Mass`
sam_body_mass = sam_body_mass_full[sam]#Find the body mass for the sample
```

Find the statistics for this sample.

```{r}
sam_mean = mean(sam_body_mass)
sam_sd = sd(sam_body_mass)

cat("Mean of the sample:", sam_mean, " Standard Deviation of the sample:", sam_sd)
```

Find the confidence interval for the mean of the full sample

```{r}
#Use the t.test method to calculate the confidence interval
sam_int = t.test(sam_body_mass, conf.level = 0.95)$conf.int

print("The confidence interval with alpha = 0.05 calcuated from the sample is shown below.")
print(sam_int)
```

#### Part C:

The interval with part c practically means that for this single sample, there is a 95% confidence that the mean of the full dataset will lie within the lower and upper bounds of the interval, [-3051.318, 23593.432].

#### Part D:

The real mean for the data is calculated below

```{r}
real_mean = mean(sam_body_mass_full)

cat("The real mean is:", real_mean)
```

This does lie in the confidence interval, and this isnt a surprise considering the 95% confidence of the interval.

#### Part E: repeat above steps for the egg length:

Find the egg lengths

```{r}
sam_egg_legnth_full = bird_data$`Egg Length`

sam_egg_legnth = sam_egg_legnth_full[sam]#Find the body mass for the sample
```

Find the statistics for this sample.

```{r}
sam_mean = mean(sam_egg_legnth)
sam_sd = sd(sam_egg_legnth)

cat("Mean of the sample:", sam_mean, " Standard Deviation of the sample:", sam_sd)
```

Find the confidence interval for the mean of the full sample

```{r}
#Use the t.test method to calculate the confidence interval
sam_int = t.test(sam_egg_legnth, conf.level = 0.95)$conf.int

print("The confidence interval with alpha = 0.05 calcuated from the sample is shown below.")
print(sam_int)
```

The interval with part c practically means that for this single sample, there is a 95% confidence that the mean of the full dataset will lie within the lower and upper bounds of the interval, [42.77990, 76.07724].

The real mean for the data is calculated below

```{r}
real_mean = mean(sam_egg_legnth_full[!is.na(sam_egg_legnth_full)])

cat("The real mean is:", real_mean)
```

This does lie in the confidence interval, and this isnt a surprise considering the 95% confidence of the interval.

#### Part F:

The bird sample was:\

|            | Number of Species Sampled | Number of Flightless Sampled |
|------------|---------------------------|------------------------------|
| Extinct    | 38                        | 21                           |
| Nonextinct | 78                        | 7                            |

Find a 95% confidence interval for the difference between the proportion of flightless birds for extinct and non-extinct species.

$$
\hat{p}_{ex} = \frac{21}{38} = 0.5526\\
\hat{p}_{non} = \frac{7}{78} = 0.0897
$$

The interval can be calculated using the following formula:

$$
int = (\hat{p}_{ex} - \hat{p}_{non}) \pm z_{\frac{\alpha}{2}}\sqrt{\frac{p_1q_1}{n_1} + \frac{p_2q_2}{n_2}}
$$

Use R as a calculator to find this confidence interval:

```{r}
#Find each catagoty of the birds:
extinct_f = 21
extinct_t = 38
living_f = 7
living_t = 78
rat_e = 21/38
rat_l = 7/78
alpha = 0.05

#The formula:
lower = (rat_e - rat_l) - qnorm(1 - alpha/2, 0, 1) * sqrt((rat_e*(1-rat_e))/extinct_t + (rat_l*(1-rat_l))/living_t)
upper = (rat_e - rat_l) + qnorm(1 - alpha/2, 0, 1) * sqrt((rat_e*(1-rat_e))/extinct_t + (rat_l*(1-rat_l))/living_t)

ci = c()
ci[1] = lower
ci[2] = upper

#For lab visualization, print them as well
cat("Lower Bound:", ci[1], "Upper Bound:", ci[2], "\n")
```

#### Part E:

The researchers findings do support this theory because the confidence interval for rat_e - rat_l is positive, meaning that the ratio of flightless to extinct birds has to be larger than to the living ones.

## 2: 7.120

#### Part A:

Find the confidence interval for the difference between the two means. Y1 is the southern pine and y2 is the ponderosa pine

$$
(\bar{y_1} - \bar{y_2}) \pm z_{\alpha / 2}\sqrt{\frac{s_{1}^{2}}{n_1} + \frac{s_{2}^{2}}{n_2}}\\
\bar{y_1} = 1312;\ \bar{y_2} = 1352;\ n_1 = 100;\ n_2 = 47;\ s_1 = 422;\ s_2 = 271\\
\\
(1312 - 1352)\pm 
$$

Calculate the z value: use normal approximation

```{r}
alpha = 0.1
z = z=qnorm(1-(alpha)/2, 0, 1)
z
```

Do the calculation:

$$
int = (1312 - 1352) \pm 1.644854 \sqrt{\frac{422^{2}}{100} + \frac{271^2}{47}}\\
$$

The interval is calculated to be [-135.1091, 55.1092].

#### Part B:

Use a computer package for this question: This will be the F test.

```{r}
var1 = 422^2
df1 = 100 - 1

var2 = 271^2
df2 = 47 - 1


#Create the confidence interval
a=.1
ci=c()
ci[1]=var1/(var2*(qf(1 - a/2, df1=df1,df2=df2)))
ci[2]=var1/var2*(qf(1 - a/2, df1=df2,df2=df1))

#For lab visualization, print them as well
cat("Lower Bound:", ci[1], "Upper Bound:", ci[2], "\n")
```

Based on these results, it is valid to say that the two shear stresses have different variances because the 90% confidence interval does not contain 1, this the interval implies that the var1 is greater than var2.

## 3: 7.128

We know that:

$$
Z = \frac{Y - \mu}{\sigma}; \ \mu = 0\\
Z =  \frac{Y - 0}{\sigma} = \frac{Y }{\sigma}\\
$$

We also know that Z^2 is distributed as a chi-sq distribution with a degree of freedom of 1

$$
Z^2 = \frac{Y^2}{\sigma^2}
$$

Because Theorum 6.11 states that:

$$
\chi^2=\frac{(n-1)S^2}{\sigma^2} 
$$

where df = (n - 1)

#### Part B:

Create a 95% confidence interval using the information from above:

#Insert IMAGE

Use R to calculate the numerical answer:

```{r}
low = 0
```

## 4: 8.24

#### Part A:

The null hypothesis is that the difference in means is 2 µm, while the alternative hypothesis would be that the mean is not 2 µm. 

#### Part B:

$$
T_{stat} = -1.02\\
p = 0.322
$$

#### Part C:

The rejection region is a mu that is less than 1.6358 and greater than 2.1262.

#### Part D: 

It can be assumed that Ho, that the the mean of the input dataset is 2, is plausible in this case.

#### Part E:

The confidence interval and the test statistic lead to the same conclusion about mu because the test statistic is derived from the confidence interval.

## 5: 8.28

#### Part A:

Import the Wisonsin lake data:

```{r}
lakes = allData$WISCLAKES

doc = lakes$DOC

meanDoc = mean(doc)
```

Use a T test to see if this is the typical for Wisconsin's lakes:

#### Calculate the t stat:

```{r}
cat("Mean concentration:", meanDoc,"\n")
cat("Number of samples:", length(doc),"\n")
cat("Std dev of data:", sd(doc))


```

$$
t = \frac{\bar y - \mu}{s/\sqrt{n}} = \frac{14.516 - 15}{12.96369/\sqrt{25}} = -0.185
$$

Find acceptable t interval

```{r}
qt(1 - 0.1/2, 25 -1 )
```

Our T value falls within [-1.71, 1.71], so we can accept the null hypothesis that the mean of the data is 15.


# ^ FIXME

```{r}
#Do the t test
t.test(doc, mu = 15, conf.level = 0.9)
```

#### Part B:

Find the prob of Ho rejection given Ho is false. This equals the prob of rejecting Ho given H1 is true. This also equals rejecting Ho given mu = 14. 

We know t of rejection = 1.711

Calculate the confidence interval: Bounds are given by

$$
y = \mu \pm t (\frac{s}{\sqrt{n}}) = 15 \pm 1.711*(\frac{12.96369}{\sqrt{25}}) = [10.564, 19.436]
$$

We reject Ho what y is outside of this interval. Now, figure out the t values for this interval:

$$
P(t < \frac{10.564 - 14}{12.963/\sqrt{25}})+ P(t > \frac{19.436 - 14}{12.963/\sqrt{25}}) = P(t < -1.325) + P(t > 2.097)
$$

Calculate this probability:

```{r}
pt(-1.325, df = 24) + (1 - pt(2.097, df = 24))
```

The probability of P(t < -1.325) + P(t > 2.097) is 0.122.

## 6: 

Read in the dataset:

```{r}
orchard = allData$ORCHARD

ratioFog = orchard$RATIO[orchard$CONDITION == "FOG"]
ratioFog = ratioFog[!is.na(ratioFog)]
ratioElse = orchard$RATIO[orchard$CONDITION == "CLEAR" | orchard$CONDITION == "CLOUD"]
```

Do a two sample t test to determine whether there is any sort of difference between the two datasets

Calculate needed values:

```{r}
v1 = var(ratioFog)
n1 = length(ratioFog)
m1 = mean(ratioFog)

v2 = var(ratioElse)
n2 = length(ratioElse)
m2 = mean(ratioElse)

v1
n1
v2
n2
m1
m2
```


$$
s^2_p = \frac{(n_1 - 1)s_1^2 + (n_2-1)s^2_2}{n_1+n_2-2} = \frac{(8 - 1)0.01407 + (4-1)0.0348}{8+4-2} = 0.02029 
$$

Calculate the test statistic:

$$
t = \frac{(\bar y_2 - \bar y _ 1) - D_o}{\sqrt{s^2_p(1/n_1+1/n_2)}} = 2.046
$$

Calculate the confidence interval

```{r}
qt(1-0.05/2, df = (n1+n2-2))
```

Since our T is between -2.22, and 2.22, it is plausible that the means of these ratios are the same between the two conditions.

## 7: 8.84

#### Part A:

Find the difference of variances: 

Read in the data:

```{r}
turbines = allData$GASTURBINE

traditional = turbines$HEATRATE[turbines$ENGINE == "Traditional"]
aero = turbines$HEATRATE[turbines$ENGINE == "Aeroderiv"]
```

Do the test using an R library: (ratio will equal 1 because vars are equal)

```{r}
var.test(traditional, aero, ratio = 1)
```

We will just use the p value for this test. Since P is way less than 0.05, we can throw out the null hypothesis that var1/var2 is 1 and conclude that the variances are different. Since the ratio. is smaller than 1, the variance of aero is larger.

#### Part B:

Repeat the above problem here with advanced turbine:

```{r}
turbines = allData$GASTURBINE

advanced = turbines$HEATRATE[turbines$ENGINE == "Advanced"]
aero = turbines$HEATRATE[turbines$ENGINE == "Aeroderiv"]
```

Do the test using an R library: (ratio will equal 1 because vars are equal)

```{r}
var.test(advanced, aero, ratio = 1)
```

This p value is very very small, meaning that the variances are almost certainly not the same. The denominator variance is much larger than the numorator. 

## 8: 8.99

#### Part A: 

Ho: The variances are the same if the ratio of var1/var2 is 1
H1: The variances are not the same if the ratio of var1/var2 is not 1.

#### Part B:

Read in the data

```{r}
ants = allData$GOBIANTS

dry = ants$AntSpecies[ants$Region == "Dry Steppe"]
gobi= ants$AntSpecies[ants$Region == "Gobi Desert"]
```

Use var test to analyze whether these samples have different variances.

```{r}
var.test(dry, gobi, ratio = 1)
```

#### Part C:

Calculate the F interval:

```{r}
qf(c(0.05/2,1-0.05/2),4,5)
```

Our F statistic is in the rejection region if it is < 0.1068 and > 7.3879

#### Part D:

The P-value of the test is 0.7264. 

#### Part E:

This means we can conclude that the variances of these two datasets are equal.

#### Part F:

We assume that these two datasets are independent.

## 9: 8.104

Read in the data for this problem:

```{r}
thru = allData$THRUPUT

hum = thru$HUMAN
aut = thru$AUTO
diff = hum - aut

meand = mean(diff)
sdd = sd(diff)
n = length(diff)

meand
sdd
n

```

Do = 0 because we think that these two methods will have the same mean

$$
t = \frac{\bar d - D_o}{s_d/\sqrt{n}} = \frac{-32.6 - 0}{35.03/\sqrt{8}} = -2.632
$$

From table 7 appendix B, we know that 
