---
title: "Lab1"
author: "Shone George Kutty Renjan"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tasks

## Task 1
Get Working Directory

```{r}
getwd()
```

## Task 2
Reads ddt.csv File

```{r}
#returns first 6 lines
ddt <-read.csv("DDT-1.csv")
head(ddt)
```

## Task 3 
Data Analysis & Subsetting


Qualitative Variables: "River" & "Species"."

Quantitative Variables: "Miles", "Length", "Width", & "ddt"



```{r}
#Number of Species
length(unique(ddt$SPECIES))
```

```{r}
# Species: LMBASS that has a WEIGHT bigger than 800gms
filtered_LMBASS_data <- ddt[ddt$SPECIES == "LMBASS" & ddt$WEIGHT > 800, ]
filtered_LMBASS_data
```

```{r}
# RIVER: SCM that has DDT bigger than 4.0
filtered_SCM_data <- ddt[ddt$RIVER == "SCM" & ddt$DDT > 4.0, ]
filtered_SCM_data

```

## Task 4: Clicker Question
### Question 1:
```{r}
# Mean length of fish in the ddt data frame
mean(ddt$LENGTH)
```

### Question 2:
```{r}
# Standard deviation of the weight of fish in the ddt data frame 
sd(ddt$WEIGHT)
```

### Question 3:
```{r}
plot(ddt$WEIGHT, ddt$LENGTH, xlab = "Weight", ylab = "Length")

# Is this the plot of LENGTH vs WEIGHT from theddt data frame?
# No, the graph is not the same because the axis are switched
```

### Question 4:
```{r}
v=1:20
v/20


# If v=1:20 what is the last value of v/20? 
# 1.00
```

## Task 5: River Tables & Barplot

Table of the rivers.
```{r}
RIVER_table= with(ddt,table(RIVER))
RIVER_table
```

Color coded barplot of the rivers.
```{r}
barplot(RIVER_table, col = c("blue","red","green", "yellow"))
```

Table of the rivers crossed with the species of fish.
```{r}
RIVER_SPECIES_table =with(ddt,table(RIVER, SPECIES))
RIVER_SPECIES_table
```
Barplot of the rivers crossed with the species of fish
```{r}
barplot(RIVER_SPECIES_table, beside=TRUE, col= c("blue","red","green", "yellow"))

```

## Task 6: Pie Chart

Fish SPECIES pie chart

```{r}
fish_speices=table(ddt$SPECIES)
pie(fish_speices, col = rainbow(3))
```


RIVER Pie Chart 

```{r}
RIV=table(ddt$RIVER)
pie(RIV,  col = c("blue","red","green", "yellow") )
```

## Task 7 : Box Plots

DDT Variable Boxplot
```{r}
boxplot(ddt$DDT)
```

WEIGHT Variable boxplot

```{r}
boxplot(ddt$WEIGHT)

```

LENGTH Variable Boxplot
```{r}
boxplot(ddt$LENGTH)

```

## Task 8: Coplots

LENGTH Vs WEIGHT given RIVER coplots
```{r}
fish_colplot=with( ddt, ifelse( SPECIES == "CCATFISH","Red",
                        ifelse(SPECIES=="SMBUFFALO","Blue","Green")))
coplot( LENGTH~WEIGHT | RIVER, data=ddt ,col=fish_colplot)
```

DDT VS WEIGTH given SPECIES Coplot

```{r}
coplot(DDT ~ WEIGHT| SPECIES, data=ddt, col = fish_colplot)
```

## Task 9: ggplot

Boxplot
```{r}
library(ggplot2)
gg_plot <- ggplot(ddt, aes(x = SPECIES, y = WEIGHT, fill = RIVER)) + geom_boxplot() + ggtitle("Shone George Kutty Renjan")
gg_plot
```


Violin Plot
```{r}
gg_plot2 <- ggplot(ddt, aes(x = RIVER, y = LENGTH, fill = SPECIES)) + geom_violin() + ggtitle("Shone George Kutty Renjan")
gg_plot2


```

Scatterplot
```{r}
gg_plot3 <- ggplot(ddt, aes(x = WEIGHT, y = LENGTH, col = SPECIES)) + geom_point() + ggtitle("Shone George Kutty Renjan")
gg_plot3


```

