---
title: "lab12"
author: "Shone George Kutty Renjan"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tasks

## Task 1: Working Directory

```{r}
getwd()
```

## Task 2: One-Sample T-Test

```{r}
set.seed(55); x1=rnorm(30, mean=25,sd=5)
```

```{r}
h22test=t.test(x1, mu=22)
h22pvalue<-h22test$p.value
h22pvalue
h22ci<-h22test$conf.int
h22ci
```

Reject $H_0: \mu = 22$ since (1) $p = 0.0021 < \alpha = 0.05$ and (2) $22 \notin CI = (23.3020, 27.2732)$$\\$

```{r}
boxplot(x1, main="Sample x1")
abline(h=c(h22ci,mean(x1)),col=c("Red","Red","Green"))

```

```{r}
h23test=t.test(x1, mu=23)
h23pvalue<-h23test$p.value
h23pvalue
h23ci<-h23test$conf.int
h23ci
```

Reject $H_0: \mu = 23$ is not possible since (1) $p = 0.0254 < \alpha = 0.05$ and (2) $23 \notin CI = (23.3020, 27.2732)$$\\$

```{r}
boxplot(x1, main="Sample x1")
abline(h=c(h23ci,mean(x1)),col=c("Red","Red","Green"))

```

```{r}
h24test=t.test(x1, mu=24)
h24pvalue<-h24test$p.value
h24pvalue
h24ci<-h24test$conf.int
h24ci
```

Accept $H_0: \mu = 24$ is possible since (1) $p = 0.1951 \geq \alpha = 0.05$ and (2) $24 \notin CI = (23.3020, 27.2732)$$\\$

```{r}
boxplot(x1, main="Sample x1")
abline(h=c(h24ci,mean(x1)),col=c("Red","Red","Green"))

```

```{r}
h25test=t.test(x1, mu=25)
h25pvalue<-h25test$p.value
h25pvalue
h25ci<-h25test$conf.int
h25ci
```

Accept $H_0: \mu = 25$ is possible since (1) $p = 0.7692 \geq \alpha = 0.05$ and (2) $25 \notin CI = (23.3020, 27.2732)$$\\$

```{r}
boxplot(x1, main="Sample x1")
abline(h=c(h25ci,mean(x1)),col=c("Red","Red","Green"))

```

```{r}
h26test=t.test(x1, mu=26)
h26pvalue<-h26test$p.value
h26pvalue
h26ci<-h26test$conf.int
h26ci
```

Accept $H_0: \mu = 26$ is possible since (1) $p = 0.4690 \geq \alpha = 0.05$ and (2) $26 \notin CI = (23.3020, 27.2732)$$\\$

```{r}
boxplot(x1, main="Sample x1")
abline(h=c(h26ci,mean(x1)),col=c("Red","Red","Green"))

```

$T_{calc}$ with $H_0=24$

```{r}
tcalc=(mean(x1)-24)/(sd(x1)/sqrt(30))
tcalc
```

```{r}
t.test(x1,mu=24)$statistic
```

```{r}
mypvalue=function(t0,xmax=4,n=20, alpha=0.05){
#calculate alpha/2
va=round(pt(-t0,df=n-1),4)
pv=2*va

# plot the t dist
curve(dt(x,df=n-1),xlim=c(-xmax,xmax),ylab="T Density",xlab=expression(t),
main=substitute(paste("P-value=", pv, " alpha=", alpha)))


# set up points on the polygon to the right
xcurve=seq(t0,xmax,length=1000)
ycurve=dt(xcurve,df=n-1)

# set up points to the left
xlcurve=seq(-t0,-xmax,length=1000)
ylcurve=dt(xcurve,df=n-1)

# Shade in the polygon defined by the line segments
polygon(c(t0,xcurve,xmax),c(0,ycurve,0),col="green")
polygon(c(-t0,xlcurve,-xmax),c(0,ylcurve,0),col="green")

# make quantiles
q=qt(1-alpha/2,n-1)
abline( v=c(q,-q),lwd=2) # plot the cut off t value 
axis(3,c(q,-q),c(expression(abs(t[alpha/2])),expression(-abs(t[alpha/2]))))


# Annotation
text(0.5*(t0+xmax),max(ycurve),substitute(paste(area, "=",va)))
text(-0.5*(t0+xmax),max(ycurve),expression(area))

return(list(q=q,pvalue=pv))
}
mypvalue(tcalc,n=30,alpha=0.05)

```

The rejection region is $(-\infty, -2.0452)\cup (2.0452,\infty)$ The p-value is 0.1952. $t_{calc}= 1.326252$ is not within the rejection region

```{r}
bootpval<-function(x,conf.level=0.95,iter=3000,mu0=0, test="two"){
n=length(x)
y=x-mean(x)+mu0  # transform the data so that it is centered at the NULL
rs.mat<-c()    #rs.mat will become a resample matrix -- now it is an empty vector
xrs.mat<-c()
for(i in 1:iter){ # for loop - the loop will go around iter times
rs.mat<-cbind(rs.mat,sample(y,n,replace=TRUE)) #sampling from y cbind -- column bind -- binds the vectors together by columns
xrs.mat<-cbind(xrs.mat,sample(x,n,replace=TRUE)) #sampling from x cbind -- column bind -- binds the vectors together by columns

}

tstat<-function(z){ # The value of t when the NULL is assumed true (xbar-muo)/z/sqrt(n)
sqrt(n)*(mean(z)-mu0)/sd(z)
}

tcalc=tstat(x) # t for the data collected
ytstat=apply(rs.mat,2,tstat) # tstat of resampled y's, ytstat is a vector and will have iter values in it
xstat=apply(xrs.mat,2,mean)  # mean of resampled x's
alpha=1-conf.level # calculating alpha
ci=quantile(xstat,c(alpha/2,1-alpha/2))# Nice way to form a confidence interval
pvalue=ifelse(test=="two",length(ytstat[ytstat>abs(tcalc) | ytstat < -abs(tcalc)])/iter,
ifelse(test=="upper",length(ytstat[ytstat>tcalc])/iter,
length(ytstat[ytstat<xstat])/iter))

h=hist(ytstat,plot=FALSE)
mid=h$mid
if(test=="two"){
ncoll=length(mid[mid<= -abs(tcalc)])
ncolr=length(mid[mid>=  abs(tcalc)])
col=c(rep("Green",ncoll),rep("Gray",length(mid)-ncoll-ncolr),rep("Green",ncolr))
}
if(test=="upper"){
ncolr=length(mid[mid>=  abs(tcalc)])
col=c(rep("Gray",length(mid)-ncolr),rep("Green",ncolr))
}

if(test=="lower"){
ncoll=length(mid[mid<=  -abs(tcalc)])
col=c(rep("Green",ncoll),rep("Gray",length(mid)-ncoll))
}
hist(ytstat,col=col,freq=FALSE,las=1,main="",xlab=expression(T[stat]))
#segments(ci[1],0,ci[2],0,lwd=2)
pround=round(pvalue,4)
title(substitute(paste(P[value],"=",pround)))
return(list(pvalue=pvalue,tcalc=tcalc,n=n,x=x,test=test,ci=ci))
}
```

```{r}
bootpval(x=x1,mu0=22,test="two")

```

```{r}
bootpval(x=x1,mu0=23,test="two")

```

```{r}
bootpval(x=x1,mu0=24,test="two")
```

```{r}
bootpval(x=x1,mu0=25,test="two")
```

```{r}
bootpval(x=x1,mu0=26,test="two")
```

The bootstrap p-value is similar to the p-value calculated using R’s t.test(). However, the bootstrap method offers an advantage by visualizing the actual T-distribution alongside the p-value area.

## Task 3: 2 Sample T-Test (Unequal Population Variance)

```{r}
set.seed(30);x=rnorm(15,mean=10,sd=7)   
set.seed(40);y=rnorm(20,mean=12,sd=4)
var.test(y,x)
```

Reject the NULL Hypothesis $H_o: \frac{\sigma_{1}^{2}}{\sigma_{2}^{2}} =1$ since (1) p-value= 0.01596 < \alpha = 0.05. (2) the 95% confidence interval (0.1048, 0.7934) is entirely below 1. Therefore, var.equal=FALSE inside t.test().

```{r}
t.test(y, x, mu =0, var.equal = FALSE)
```


```{r}
t.test(y, x, mu =2, var.equal = FALSE)
```
The Welch 2 Sample T-Test's Degrees of Freedoms is not an integer. The confidence interval is wider than it would be if the population variance are equal> The result is that we can't reject either NULL Hypotheses not being possible.

## Task 4:  2 Sample T-Test (Equal Population Variance)

```{r}
set.seed(30);x2=rnorm(15,mean=10,sd=4)   
set.seed(40);y2=rnorm(20,mean=12,sd=4)
var.test(y2,x2)
```
Accept the NULL Hypotheses that the True  Ratio of Variance Equal to 1 as pasuible. Therefore assign var.equal=TRUE inside t.test() as follows

```{r}
t.test(y2,x2,mu=0, var.equal = TRUE)
```

```{r}
t.test(y2,x2,mu=2, var.equal = TRUE)

```
Assuming that the population variances are equal leads to a smaller confidence interval, which allows us to also reject the first null hypothesis (null). value = 0)It appears that the standard error is not possible. It seems to be a relevant factor in our Hypothesis Testing Procedure.


## Task 5: boot2pval() with Task 3
```{r}
boot2pval<-function(x1,x2,conf.level=0.95,iter=3000,mudiff=0, test="two"){
n1=length(x1)
n2=length(x2)
y1=x1-mean(x1)+mean(c(x1,x2))  # transform the data so that it is centered at the NULL
y2=x2-mean(x2)+mean(c(x1,x2))
y1rs.mat<-c()    #rs.mat will be come a resample matrix -- now it is an empty vector
x1rs.mat<-c()
y2rs.mat<-c()
x2rs.mat<-c()
for(i in 1:iter){ # for loop - the loop will go around iter times
y1rs.mat<-cbind(y1rs.mat,sample(y1,n1,replace=TRUE)) #sampling from y cbind -- column bind -- binds the vectors together by columns
y2rs.mat<-cbind(y2rs.mat,sample(y2,n2,replace=TRUE))

}
x1rs.mat<-y1rs.mat+mean(x1)-mean(c(x1,x2))
x2rs.mat<-y2rs.mat+mean(x2)-mean(c(x1,x2))

xbar1=mean(x1)
xbar2=mean(x2)
sx1sq=var(x1)
sx2sq=var(x2)

tcalc=(xbar1-xbar2-mudiff)/sqrt(sx1sq/n1+sx2sq/n2)

sy1sq=apply(y1rs.mat,2,var)
sy2sq=apply(y2rs.mat,2,var) 
y1bar=apply(y1rs.mat,2,mean)
y2bar=apply(y2rs.mat,2,mean)

tstat=(y1bar-y2bar-mudiff)/sqrt(sy1sq/n1+sy2sq/n2)


alpha=1-conf.level # calculating alpha
#ci=quantile(xstat,c(alpha/2,1-alpha/2))# Nice way to form a confidence interval
pvalue=ifelse(test=="two",length(tstat[tstat>abs(tcalc) | tstat < -abs(tcalc)])/iter,
ifelse(test=="upper",length(tstat[tstat>tcalc])/iter,
length(ytstat[tstat<tcalc])/iter))

h=hist(tstat,plot=FALSE)
mid=h$mid
if(test=="two"){
ncoll=length(mid[mid<= -abs(tcalc)])
ncolr=length(mid[mid>=  abs(tcalc)])
col=c(rep("Green",ncoll),rep("Gray",length(mid)-ncoll-ncolr),rep("Green",ncolr))
}
hist(tstat,col=col,freq=FALSE)
#segments(ci[1],0,ci[2],0,lwd=2)

return(list(pvalue=pvalue))
#return(list(pvalue=pvalue,tcalc=tcalc,n=n,x=x,test=test,ci=ci))
}
set.seed(30);x=rnorm(15,mean=10,sd=7)   
set.seed(40);y=rnorm(20,mean=12,sd=4)
```

```{r}

boot2pval(x1=y,x2=x)
```
```{r}
boot2pval(x1=y,x2=x,mudiff=2)

```

## Task 6: boot2val() with Task 4

```{r}
set.seed(30);x2=rnorm(15,mean=10,sd=4)   
set.seed(40);y2=rnorm(20,mean=12,sd=4)
boot2pval(x1=y2, x2=x2)
```

```{r}
boot2pval(x1=y2, x2=x2,mudiff = 2)

```

## Task 7: Explain t.test() output

Line A: performs a one-sample T test for Null Hypothesis:$H_o : \mu =23$  using x as the sample.$\\$
Line B: The title of the results of the T-test indicates that it was conducted on a single sample and the name of the test according to R.$\\$
Line C: provides the $t_{calc} =2.3563$ with 29 degrees of freedom (also imples sample size of 30), p-value=0.02543 represents the area below minus tcalc and above positive tcalc.$\\$
Line D : states the Alternative Hypothesis( The Researcher's Belief) as $H_o : \mu \neq 23$.$\\$
Line E: Shows the confidence level, below this line is the actual interval.$\\$
Line F: 95% confidence interval for $\mu$ is (23.3020, 27.2732).$\\$
Line G: The point estimate for the true mean using the mean x1 is 24.28759.$\\$

## Task 8: function Package
```{r}
set.seed(55); x1=rnorm(30,mean = 25,sd=5)
MATH4753SHON2024::bootpval(x1,mu=25)
```

