---
title: "Lab 5"
author: "Shone George Kutty Renjan"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

# Tasks

## Task 1

```{r}
getwd()
```

## Task 2
```{r}
mybin=function(iter=100,n=10, p=0.5){ 
# make a matrix to hold the samples
#initially filled with NA's
sam.mat=matrix(NA,nr=n,nc=iter, byrow=TRUE)
#Make a vector to hold the number of successes in each trial
succ=c()
for( i in 1:iter){
#Fill each column with a new sample
sam.mat[,i]=sample(c(1,0),n,replace=TRUE, prob=c(p,1-p))
#Calculate a statistic from the sample (this case it is the sum)
succ[i]=sum(sam.mat[,i])
}
#Make a table of successes
succ.tab=table(factor(succ,levels=0:n))
#Make a barplot of the proportions
barplot(succ.tab/(iter), col=rainbow(n+1), main="Binomial simulation", xlab="Number of successes")
succ.tab/iter
}

```

Record the plots for each of the following 

### 100 iteration
```{r}
mybin(iter=100, n=10, p=0.7)

```
### 200 iteration
```{r}
mybin(iter=200, n=10, p=0.7)

```


### 500 iteration
```{r}
mybin(iter=500, n=10, p=0.7)

```

### 1000 iteration
```{r}
mybin(iter=1000, n=10, p=0.7)

```

### 10000 iteration & Table
```{r}
mybin(iter=10000, n=10, p=0.7)

```
### Check value aproximately Using Dbinom
```{r} 
dbin.tab<-round(dbinom(0:10, size = 10, prob = 0.7), 4)
names (dbin.tab) = 0:10
dbin.tab


```
The dbinom and the 10000 iteration are about the same and aproximately correct.

## Task 3
### Sample Funtion of n=5 without Replacement 
```{r}
sample (c(1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0), size = 5, replace = FALSE, prob = NULL)
```

### Sample Funtion of n=5 with Replacement 

```{r}
sample(c(1,1,1,1,1,1,1,1,1,1,1,1,0,0,0,0,0,0,0,0), size =5,replace = TRUE, prob = NULL)

```

### myhper Function

```{r}
myhyper=function(iter=100,N=20,r=12,n=5){
# make a matrix to hold the samples
#initially filled with NA's
sam.mat=matrix(NA,nr=n,nc=iter, byrow=TRUE)
#Make a vector to hold the number of successes over the trials
succ=c()
for( i in 1:iter){
#Fill each column with a new sample
sam.mat[,i]=sample(rep(c(1,0),c(r,N-r)),n,replace=FALSE)
#Calculate a statistic from the sample (this case it is the sum)
succ[i]=sum(sam.mat[,i])
}
#Make a table of successes
succ.tab=table(factor(succ,levels=0:n))
#Make a barplot of the proportions
barplot(succ.tab/(iter), col=rainbow(n+1), main="HYPERGEOMETRIC simulation", xlab="Number of successes")
succ.tab/iter
}
```

### 100 iteration
```{r}
myhyper(iter=100,N=20,r=12,n=5) 

```
### 200 iteration
```{r}
myhyper(iter=200,N=20,r=12,n=5) 

```

### 500 iteration
```{r}
myhyper(iter=500,N=20,r=12,n=5) 

```

### 1000 iteration
```{r}
myhyper(iter=1000,N=20,r=12,n=5) 

```


### 10000 iteration & table
```{r}
myhyper(iter=10000,N=20,r=12,n=5) 

```

### Check value aproximately Using Dhyper
```{r}
dhyper.tab<-round(dhyper(x= 0:5, m=12, n=8,k=5), 4)
names (dhyper.tab) = 0:5
dhyper.tab

```
The 10000 iteration and the dhyper value are almost the same and aproximately correct


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Tasks 4 
### What does the mysample() function do?

The mysample function creates random samples of a specified size by choosing numbers from 1 to 10 repeatedly. It visualizes the results using a bar graph that displays how often each number appears. The function allows for adjusting the pause duration between iterations. In each cycle, it generates a new sample (allowing repeat selections), tallies the occurrences, and updates the graph, with a customizable delay between updates.


### mysample(n=1000, iter=30,time=1)
 Display the graph one time
```{r}
mysample=function(n, iter=30, time=0.5){
  for(i in 1:iter){
    # make a sample
    s = sample(1:10, n, replace = TRUE)
    # turn the sample into a factor
    sf = factor(s, levels = 1:10)
    # make a barplot for the first iteration only
    if(i == 1) {
      barplot(table(sf)/n, beside = TRUE, col = rainbow(10),
              main = paste("Example sample()", " iteration ", i, " n= ", n, sep = ""),
              ylim = c(0, 0.2)
      )
    }
    # delay for visualization purposes
    Sys.sleep(time)
  }
}

# Run the function
mysample(n = 1000, iter = 30, time = 1)

```

### Describe what you see here.
30 bar graphs representing the frequencies of values from 1 to 10, where some bars are noticeably taller than others. This illustrates that, although we'd expect the bars to be of equal height, random independent selection can lead to certain values being chosen more frequently. This demonstrates that, to achieve more similar frequencies, a larger sample size (n) is needed, as it would result in a more even distribution.


## Task 5
### Choose(8,5)
```{r}
choose(8,4)
```

$$
(^8 _5)=70
$$

```{r}
dnbinom(10,size=3,prob=0.4)
```

### more calculations in R
```{r}
pbinom(q=8,size=15,prob = 0.4)
```


## Task 6
```{r}
MATH4753SHON2024::mybin()
```
