---
title: "Lab2"
author: "Shone George Kutty Renjan"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tasks

## Task 1
```{r}
getwd()
```

## Tasks 2
```{r}
epagas=read.csv("EPAGAS.csv")
head(epagas)
```

## Task 3
Make the object mpg, the number of miles per gallon vector.
```{r}
mpg=epagas$MPG
head(mpg)
```

Transform the mpg variable to z and verify these results.
```{r}
z=(mpg -mean(mpg)) /sd(mpg)
head(z)
print(round(mean(z), 4))

# Mean of z-score function 
print(paste("Mean of z-score = ", round(mean(z), 4)),quote = FALSE)
# Another way to find the mean of z-score
#print(round(mean(z), 4))

# Variance of z function
print(paste("Variance = ", round(var(z), 4)), quote= FALSE)
# Another way to get the Variance for z
#sd(z)**2

# Standard Deviation of z function
print(paste("Standard Deviation = ", round(sd(z), 4)), quote= FALSE)
```
Values of mpg possible outliers.
```{r}
mpg[abs(z) >= 2 & abs(z) <= 3]
```

Values of mpg that defined as outliers
```{r}
mpg[abs(z) > 3]
```


Construct a dotplot with colors, Red=outlier, Blue=possible outlier
```{r}
outlier_dotp= ifelse(abs(z)>3, "Red",
                     ifelse(abs(z)>= 2  & abs(z) <= 3, "Blue", "Black"))

library(lattice)
dotplot(mpg, col=outlier_dotp)
```


## Task 4
Make a boxplot of the mpg variable
```{r}
boxplot(mpg, col= "black", notch = TRUE, main ="MPG BOXPLOT", horizontal =TRUE, xlab= "MPG" )
```


### Chebyshev’s theorem
#### 1-(1/k^2)

Proportion of data within 2 standard deviations of the mean of the mpg data o	using 
```{r}
k=2
sd=1-(1/k^2)
print(sd)
```

Proportion within 2 standard deviation of the mean
```{r}
two_sd =length(mpg[abs(z)<2])
two_sd/length(mpg)

```

### Chebyshev's Rule
Chebyshev's Rule is valid for this data, because it say that at least 75% are between the two deviation of the mean

### Empirical Rule
The Empirical Rule applies to data that is unimodal and roughly symmetrical around the mean, with about 95% of the data falling within two standard deviations. Since data was 96% within this range, your data is close to this rule. To verify, check if the data distribution has a single peak (unimodal) and is symmetrical around the mean.


```{r}
density_mpg = density(mpg)
plot(density_mpg)
abline(v=mean(mpg), col = "red")
```

The Empirical Rule is valid because it is (unimodal) and is symmetrical around the mean.
